|

### 2024-09-06 manuscript/paperCMSACversion.Rmd
Most recent updates.

@@

While we found this work to be of interest, we believe that a more appropriate statistical framework for this analysis is survival analysis.  Therefore, we performed our own extended analysis with additional data using methods from survival analysis to estimate difference survival times between enforcers and non-enforcers.  For enforcers, defined based on number of career fights we find no significant difference in survival times.  Conversely, we do find a significant difference in survival times for enforcers defined based on average penalty minutes per game.  Specifically, we find that enforcers defined in this way have a hazard ratio of 2.34988

Unobserved Confounders: 
We further note several additional difference between the enforcers and the non-enforcer groups that were not considered in the @Popkin2023.  As an example, the number of average number of goals scores between the two groups significantly different for both definitions of enforcers.  The average number of goals scores for E-F enforcers  vs non-enforcers was `r summ$mngoals[summ$EF]` vs `r summ$mngoals[!summ$EF]` (p-value: `r ifelse(round(t.test(dat$G~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EF)$p.value,2))`).  For E-P enforcers, the average number of goals for enforcers was `r summEP$mngoals[summEP$EP]` vs `r summEP$mngoals[!summEP$EP]` (p-value: `r ifelse(round(t.test(dat$G~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EP)$p.value,2))`).

@@

Perhaps more interestingly, we find highly significant differences in geographic region of enforcers vs non-enforcers for both definitions.  In the enforcer group `r paste0(100*round(efcountry$p[efcountry$EF],4),"\\%")` and `r paste0(100*round(epcountry$p[epcountry$EP],4),"\\%")` of those groups are from North America (i.e. United States or Canada) for E-F and E-P, respectively.  For non-enforcers, North American players make up only `r paste0(100*round(efcountry$p[!efcountry$EF],4),"\\%")` for E-F and `r paste0(100*round(epcountry$p[!epcountry$EP],4),"\\%")` for E-P.  For both definitions of enforcer, these differences are significant with p-values for both definitions less than 0.0001.

While we do not address this imbalance here, as we focus on closely following the work of @Popkin2023, it is worth noting the substantial imbalance in geographic origin (i.e. North American vs outside of North America) of enforcers versus non-enforcer.

>>>>

We note here that figure \ref{fig:KMEP} demonstrates clearly why mortality rate is often not an appropriate measure for comparing risk between groups.  If one were to look only at mortality rates at age 75, the mortality rates among these two groups are nearly identical (just under 20\%).  However, the enforcers survival curve is clearly below the non-enforcer survival curve estimate at almost every point below 75 indicating that enforcers are dying at younger ages on average even though the mortality rate eventually becomes nearly indistinguishable near the age of 75.  All of this is, by the way, exactly consistent with the findings of @Popkin2023 (i.e. no signficant difference in mortality rates, but enforcers dying at younger ages).

@@

@@

While we found this work to be of interest, we believe that a more appropriate statistical framework for this analysis is survival analysis.  Therefore, we performed our own extended analysis with additional data using methods from survival analysis to estimate difference survival times between enforcers and non-enforcers.  For enforcers, defined based on number of career fights we find no significant difference in survival times.  Conversely, we do find a significant difference in survival times for enforcers defined based on average penalty minutes per game.  Specifically, we find that enforcers defined in this way have a hazard ratio of 2.34988 (p-value: 0.00125) relative to non-enforcers indicated that there are significant risks associated with the position of enforcer.  
While we find this clear association between enforcers and increased hazard by controlling for the variables that were matched on in @Popkin2023, there are several obvious uncontrolled variables that have substantially different distributions in the enforcer and non-enforcer populations.  One simple example are the differences in average number of goals scores between the two groups significantly different for both definitions of enforcers.  The average number of goals scores for E-F enforcers  vs non-enforcers was `r summ$mngoals[summ$EF]` vs `r summ$mngoals[!summ$EF]` (p-value: `r ifelse(round(t.test(dat$G~dat$EF)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EF)$p.value,2))`).  For E-P enforcers, the average number of goals for enforcers was `r summEP$mngoals[summEP$EP]` vs `r summEP$mngoals[!summEP$EP]` (p-value: `r ifelse(round(t.test(dat$G~dat$EP)$p.value,2) <0.0001,"<0.0001",round(t.test(dat$G~dat$EP)$p.value,2))`).

Perhaps more interestingly, we find highly significant differences in geographic region of enforcers vs non-enforcers for both definitions.  In the enforcer group `r paste0(100*round(efcountry$p[efcountry$EF],4),"\\%")` and `r paste0(100*round(epcountry$p[epcountry$EP],4),"\\%")` of those groups are from North America (i.e. United States or Canada) for E-F and E-P, respectively.  For non-enforcers, North American players make up only `r paste0(100*round(efcountry$p[!efcountry$EF],4),"\\%")` for E-F and `r paste0(100*round(epcountry$p[!epcountry$EP],4),"\\%")` for E-P.  For both definitions of enforcer, these differences are significant with p-values for both definitions less than 0.0001.  While we do not address this imbalance here, as we focus on closely following the work of @Popkin2023, it is worth noting this imbalance in geographic origin (i.e. North American vs outside of North America) of enforcers versus non-enforcer, and we leave the exploration of this issue for future work.
